---
title: "Initial Korean covid 19"
author: "Andrew Valles Tovar"
date: "3/30/2023"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?






## Graphical results - use ggplot commands to do this  

```{r}
barchartGC(~result + sex,data=kRdata)
barchartGC(~result + sex,data=kRdata, type="percent")
```

According to the data, it seems more percentage of males have died more than females. On the other side of the chart, females are more alive than males are.


  
## Numerical results
```{r}
table1 <- xtabs(~result + sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

The row data shows the amount of females who caught COVID that are alive is 56% and 44% of the males are alive. However; out of the ones that died to the disease, 64% of them were male and 36% were female.

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
In Person's chi-squared test, we can see the total amount of people from each gender that caught COVID. 1817 females who caught the disease are alive, but 25 of them are dead. 44 males who caught the disease are deceased, but 1425 of them are alive. The counts expected by null show that slightly more females died to the disease and males died slightly less to it. The odds ratio is 2.3, which means males are more likely to die from COVID than females.

## Conclusions
In summary, males died more to COVID-19 than females in Korea. Although the data shows only a small amount of people died to the disease, it was males that died the most than compared to females. The inferential data shows that the total amount of females that caught the disease were 1817, which is more than males by 392.So it's surprising that even with a bigger pool of females, males are the one that died the most. Males lead the death count by 44 and females have a death count of 25.
